### 2017-12-20 paper/paper.tex

\documentclass{article}

\usepackage{float}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage[hyperref=true, natbib=true, style=numeric, backend=bibtex]{biblatex}
\usepackage{graphicx}
\bibliography{references.bib}

\title{TCP Congestion Algorithms In Datacenters \\
	\vspace{0.3cm}
	{\large LS Project}
}
\date{\today{}}
\author{Luc Gommans, Rick van Gorp}

\begin{document}
\maketitle

\section{Abstract}

In datacenters, multiple tenants share a finite amount of bandwidth. When an
uplink is saturated, packets that overflow available buffer space will be
dropped and have to be retransmitted by the sender. When one tenant chooses to
retransmit faster than others, they have an advantage as they will make it
though more quickly. TCP is a connection-oriented protocol which employs
congestion control algorithms. These algorithms aim to divide the available
bandwidth equally between network flows.

Some congestion control algorithms are more aggressive than others, and it has
been reported that some algorithms cause others to back off excessively.
Tenants may, perhaps inadvertedly, cause inequal bandwidth distributions during
times of congestion.

In this research, we compare five algorithms under different network conditions
to assess how they influence each other. We found that only CTCP (Windows)
backs off excessively; all other algorithms behave fairly. The results for CTCP
are anomalous compared to all other algorithms and we conclude that for this
algorithm, future work is needed for a reliable assessment.

\section{Introduction}

Multitentant datacenters aim to provide services with consistent performance
and reliability. When customers send a lot of data in a short amount of time,
networks might become saturated or even congested, and bandwidth needs to be
distributed fairly among tenants.

Congestion occurs when a buffer on the path to the recipient is full and
packets have to be dropped. TCP uses a congestion control algorithm to avoid
congestion in the network and distribute bandwidth fairly among users. Several
algorithms are available to perform this function. Each has a different behaviour
depending on the network's characteristics.

This report includes a performance and behavior analysis of different TCP congestion control algorithms in a (simulated) multitenant datacenter network. Based on the results, we will recommend datacenters on measures to take to keep providing a reliable environment for customers when multiple congestion control algoritms are active in the network, for example using Quality of Service or discouraging certain congestion control algorithms. The main research question is: {\it How can a datacenter provide a reliable and performant environment for customers when multiple TCP congestion control algorithms are used in the
network?}

In the next section, we will discuss related work. Section \ref{sec:method}
describes our methods: how we selected algorithms to test, how we tested their
performance, and how we chose network conditions to simulate. Section
\ref{sec:results} describes the results of our research: the algorithms we
chose to test are described in \ref{sub:chosen-algos}, a theoretical analysis
of those is described in \ref{sub:algos-theoretical}, and experimental results
can be found in \ref{sub:experiment-results}. We discuss our results in section
\ref{sec:discussion} and draw conclusions in section \ref{sec:conclusion}.

\section{Related work}

In 2010, the algorithm DCTCP was described by Mohammad Alizadeh et
al.\cite{dctcp-congestion-original} and published as
RFC8257\cite{dctcp-congestion} in 2017. DCTCP is optimized for datacenters and
provides a high-burst tolerance, low latency, and high throughput when the
datacenter has a small part of the buffer available\cite{dctcp-congestion}.

In 2016 another algorithm was proposed: BBR. This is a TCP congestion algorithm
created by Google, which achieves higher bandwidths and lower latencies
compared to other TCP congestion methods\cite{bbr-congestion}. A comparison of
BBR with CUBIC\cite{bbr-congestion-comparison} shows that the BBR node pushes
the CUBIC node away in bandwidth when using small buffers. The BBR node gets
more bandwidth allocated than the CUBIC node.

The TCP congestion control algorithm BBR is discussed by Neal Cardwell et
al.\cite{bbr-congestion} The article contains a performance test of the BBR
algorithm and a comparison between the BBR algorithm and the CUBIC algorithm,
where a noticeable difference in bandwidth allocation is shown when using small
buffers.

A. Esterhuizen and A.E. Krzesinski\cite{multiple-congestion} analyse the
performance of multiple TCP congestion control algorithms, specifically: Reno,
BIC, CUBIC, HighSpeed TCP, TCP-Hybla, TCP-Illinois, TCP Low Priority,
TCP-Vegas, TCP-Westwood, TCP-YeAH, and Scalable TCP.

\section{Method}\label{sec:method}

During this research, performance comparison tests of various TCP congestion algorithms will be performed. The most common TCP congestion algorithms will be determined through desk research. The tests will be conducted in an isolated environment. Based on the results of those tests, recommendations will be given related to the measures a datacenter has to take to keep the traffic equally allocated over customers.

\subsection{Selecting TCP congestion algorithms}

In order to select the most common TCP congestion algorithms desk research will be performed. This desk research will show which algorithms are used more often than other algoritms by looking for the default algorithms in Operating Systems. Based on those results a list of TCP congestion algorithms will be composed.

\subsection{Comparing Performance}

We will analyse one common scenario: multiple servers of different owners
inside a multitenant datacenter, sends data to clients outside the datacenter.
This could be users downloading software, the bursts while streaming a video
or movie, or large resources on a webpage.

We will create a bottleneck on the path to the user to see how different
congestion algorithms handle this. The connection will be full duplex and have
symmetrical speeds, to avoid any congestion in sending acknowledgements.
Because the server is sending the large volume of data and the client is
responding with the much smaller acknowledgements, the server will be the one
applying the congestion control. Therefore we vary the algorithm settings of
the server and not the client. The client will always be a GNU/Linux machine
using CUBIC, but we assume (due to the aforementioned reasons) that this does
not matter.

The bottleneck is created by having one receiving host, listening on different
ports, and two transmitting hosts. Each network interface is 1 gigabit per
second, thus the servers will have to limit their speed in order to avoid
flooding the client. This setup is shown in figure \ref{fig:setup1}.

We use three physical hosts as servers: two running Ubuntu 17.04 and one
running Microsoft Windows 10. In each experiment, only two hosts will be used.
The Windows system is used because there is no working CTCP
implementation---the default algorithm in Windows since Vista---available for a
modern Linux kernel.

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.5]{figs/setup2.png}
		\caption{Test setup - Scenario 1}
	\label{fig:setup1}
\end{figure}

We will test the TCP congestion algorithm for fairness in sharing bandwidth
with other TCP connections using other congestion algorithms. In order to test
the fairness, the bandwidth usage per second per algorithm is measured. If two
different algorithms use similar amounts of bandwidth, they can be said to be
fair relative to each other.

The tests conducted with the algorithms have two environment variables: packet loss and delay. Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}. \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay. Windows, which is tested as a server, does not support \texttt{netem}. Therefore the packet loss and delay are applied to the incoming interfaces of the clients. Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [TODO: SRC]. The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

Each test is run at least twice. When results are conflicting or unexpected, we
ran a test more often and on different hardware.

%\item TCP Window Size measured over time: to determine whether using a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of a larger window or a smaller window to a specific TCP connection.
% \item Round Trip Time: to determine the reason of the behaviour of a TCP congestion algorithm in combination with other TCP congestion algorithms.

\subsubsection{Determining the Packet loss and Delay}

The dataset of Verizon \cite{verizon-latency} is used to calculate the packet loss and delay used in this research. The dataset includes global information and shows averages per month over the year 2017. The packet loss is calculated by taking the average of the Verizon Business Packet Delivery Statistics for Country Specific Metrics for each country. Those averages were inverted by subtracting the packet delivery percentage from 100, to show the actual packet loss in percentages and not the packet delivery. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:packet-loss-chart}

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.7]{figs/verizon-packetloss.png}
		\caption{Average Packet Loss for Country Specific Metrics}
	\label{fig:packet-loss-chart}
\end{figure}

The country numbers in figure \ref{fig:packet-loss-chart} are only used to show the distribution of the average packet loss. The maximum packet loss shown in figure \ref{fig:packet-loss-chart} equals 0.6\% and the minimum equals 0\%. However, in case of situations where more packet loss occurs than usual, the double value of the maximum is used: 1.2\%. The values for packet loss in this comparison set-up were determined based on the distribution shown in figure \ref{fig:packet-loss-chart} and the manually set maximum of 1.2\%. The values are shown in table \ref{table:test-packetloss}.

\begin{table}[H]
	\centering
	\caption{Packet loss percentages used for the comparison setup}
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Packet loss} \\
	\hline 0\%. \\
	\hline 0.01\% \\
	\hline 0.1\% \\
	\hline 0.6\% \\
	\hline 1.2\% \\
	\hline
	\end{tabular}
	\label{table:test-packetloss}
\end{table}

The delay is calculated by taking the average of the Verizon Business Latency Statistics for Country Specific Metrics for each country. Based on the ascending sorted values a bar chart was created. This bar chart is shown in figure \ref{fig:verizon-delay-chart}

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.7]{figs/verizon-delay.png}
		\caption{Average Latency for Country Specific Metrics}
	\label{fig:verizon-delay-chart}
\end{figure}

The country numbers in figure \ref{fig:verizon-delay-chart} are only used to show the distribution of the average delays. The maximum delay shown in figure \ref{fig:verizon-delay-chart} equals 290ms and the minimum equals 8ms (both rounded). The values for delay in this comparison set-up were determined based on the distribution shown in figure \ref{fig:verizon-delay-chart}. The values are shown in table \ref{table:test-delay}.

\begin{table}[H]
	\centering
	\caption{Packet loss percentages used for the comparison setup}
	\begin{tabular}[H]{ | l |}
	\hline
	\textbf{Delay} \\
	\hline  8ms\\
	\hline  64ms  \\
	\hline  120ms  \\
	\hline  176ms \\
	\hline	232ms \\
	\hline  290ms  \\
	\hline
	\end{tabular}
	\label{table:test-delay}
\end{table}

\subsubsection{Performing time-based automated tests}

The tests will be performed using scripts written in Python [\ref{appendix:python}]. The scripts either act as controller or as disciple. The controller is responsible for controlling the tests, by synchronizing the time, exchanging commands to run, modify or stop the tests and gathering data. The disciple is the listener and executes all task commands the controller sends. The commands include getting the hostname of a disciple, setting the TCP congestion algorithm of a disciple, configuring the time of a disciple, starting the data transfer and measurement tools on the disciple and closing the connection with the controller.

\subsubsection{Testing hardware}

The specifications of the hardware used for the tests are shown in table \ref{table:spec1}, table \ref{table:spec2} and table \ref{table:spec3}. The switch shown in table \ref{table:spec3} has all router related functionalities, which might impact the network speed, turned off.

\begin{table}[H]
	\centering
	\caption{Node configurations}
	\begin{tabular}[H]{ | l | l | }
	\hline
	\textbf{Hardware} & \textbf{Specification} \\
	\hline  Model & Dell Optiplex 7010\\
	\hline  CPU & Intel(R) Core(TM) i5-3570S CPU @ 3.10GHz\\
	\hline  Memory & 12GB\\
	\hline  NIC & Intel Corporation 82579LM Gigabit Network Connection (rev 04)\\
	\hline	NIC Driver & e100e 3.2.6-k\\
	\hline
	\end{tabular}
	\label{table:spec1}
\end{table}

\begin{table}[H]
	\centering
	\caption{Switch configuration}
	\begin{tabular}[H]{ | l | l | }
	\hline
	\textbf{Hardware} & \textbf{Specification} \\
	\hline  Model & 3Com OfficeConnect Gigabit Switch 16\\
	\hline  Ports & 16x Gigabit Ethernet\\
	\hline
	\end{tabular}
	\label{table:spec2}
\end{table}

\subsection{Traffic Management}

Desk research will be conducted to gather information related to traffic management. The goal is to find a mechanism that is capable of allocating bandwidth equally between customers of a datacenter. This is done by analyzing the results from the algorithm comparison and the initial desk research on traffic management.

\section{Results}\label{sec:results}

In this section we will first elaborate on which algorithms we looked into, next we look at their internal workings, and finally we show the results of our experiments.

\subsection{TCP Congestion Algorithms}\label{sub:chosen-algos}

The most common TCP congestion algorithms are CTCP, CUBIC and BIC. CTCP is the default congestion algorithm in Microsoft Windows Server 2008 and newer \cite{cubic-kernel-version}. CUBIC is the default congestion algorithm in Linux distributions with kernel versions 2.6.19 and higher \cite{cubic-kernel-version}. BIC was the default congestion algorithm in earlier Linux distributions from kernel version 2.6.8 until 2.6.19 \cite{bic-kernel-version} \cite{cubic-kernel-version}. DCTCP and BBR will also be tested as those protocols are new and upcoming---they are included in Linux since 4.9 (December 2016) \cite{linux-bbr}. DCTCP is specifically designed for datacenters \cite{dctcp-congestion} and BBR is a new algorithm meant for general use \cite{bbr-congestion}.

\subsection{Theoretical Algorithm Characteristics}\label{sub:algos-theoretical}

The BBR algorithm depends on parameters $BtlBw$ and $RT_{prop}$ \cite{bbr-congestion}. Those parameters are estimates defined by the BBR algorithm. $BtlBw$ specifies the bandwidth at the slowest link in each direction and $RT_{prop}$ is the round-trip propagation time. $RT_{prop} = RTT - queuing delay - processing delay$, which results in the minimum amount of time for round trip propagation in case there are no queueing or processing delays. Using those two parameters, the Bandwidth Delay Product (BDP) is calculated, which is the maximum possible amount of data being sent in a network: $BDP = BtlBw * RT_{prop}$. As long as the amount of data in transit is less than the $BDP$ and $BtlBw$, the delivery rate is increased. If the amount of data in transit is equal to $BtlBw$ the delivery rate can not go up anymore. The $RTT$ can never be lower than $RT_{prop}$... explain more on delay perhaps?.

The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}. For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced. For the delay-based component, the variable delay window, $dwnd$, was introduced. The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver. For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$. The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active. $dwnd$ is calculated differently for several scenarios. If the network path is underutilized: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$. When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$. When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$. $k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimization of the algorithm. $\eta$ defines the rapidness of reduction of the window when congestion is detected. $\beta$ is the factor used for the multiplicative decrease after loss is detected. $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).

The DCTCP algorithm consists of three components \cite{dctcp-congestion-original}:

\begin{enumerate}
	\item Simple Marking at the Switch. Based on the marking treshold $K$ a packet is marked with the CE flag. This only occurs when the queue occupancy is greater than K at arrival time. The marking treshold $K$ specifies the minimum value of queue occupancy.
	\item ECN-Echo at receiver. The receiver running the DCTCP algorithm ACKs every packet and sets the ECN-Echo flag only if the packet was marked with CE. The ECN-Echo flag is used to notify the sender of congestion in the network.
	\item Controller at the sender. The sender has an estimate of the fraction of packets that are marked with CE. This is based on the actual fraction of packets that were marked, the weight given to the new estimate compared to the past weight. This results in $\alpha$. $\alpha$ is used to indicate the congestion on a network with a maximum of value 1, indicating a congested network, and a minimum value of 0, indicating no congestion in the network. DCTCP cuts the window size using $\alpha$: $cwnd = cwnd * (1 - \alpha/2)$. This results in a low queue length, but ensuring high throughput.\\
\end{enumerate}

The BIC algorithm controls its windows based on the size of windows \cite{bic-tcp-congestion}. It consists of two components:

\begin{enumerate}
	\item Binary search increase. This algorithm computes the midpoint between the maximum window size $W_{max}$ and current minimum window size $W_{min}$. If packet loss occurs, $W_{max}$ is set to the midpoint. This process repeats until $W_{max}$ and $W_{min}$ are smaller than the minimum increment $S_{min}$.
	\item Additive increase runs as an addition to the binary search increase. If $W_{max} - midpoint > S_{max}$, the window size is increased by $S_{max}$ until $W_{max} - midpoint < S_{max}$. $S_{max}$ is the preset maximum increment.
\end{enumerate}
In case the current window size becomes greater than $W_{max}$ a slow start strategy is performed to determine a new $W_{max}$: $W_{max} + \alpha * S_{min}$, where $\alpha$ is a multiplier. The multiplier is increased by one until $S_{min} * \alpha >= S_{max}$ or when losses occur. In the case of loss a new value for $W_{max}$ will be set.

The CUBIC algorithm is a successor of the BIC algorithm \cite{cubic-tcp-congestion}. In case of a loss event $W_{max}$, which is the maximum window size, is registered. This is followed by a multiplicative decrease of the congestion window by factor $\beta$. If an ACK in congestion avoidance is received and the current window size $cwnd$ is less than $W_{max}$, the algorithm uses the concave profile to increase the window size according to the formula: $cwnd = \frac{W(t+RTT)-cwnd}{cwnd}$. If $cwnd$ is larger than $W_{max}$, the convex profile is used to increase the window size according to the same formula as the concave profile. In both formulas $t$ is the time elapsed from the last window reducation.

\subsection{Experiments' Results}\label{sub:experiment-results}

In our results, we see that all algorithms clearly respond to the simulated
network circumstances, but not very much to each other. Most algorithms are
fair towards all other algorithms and will use their half of the connection.
See figures \ref{fig:fair-1} and \ref{fig:fair-2}.

\begin{figure}[H]
	\makebox[\textwidth]{
		\centering
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/cubic-bic-8-0.png}
			\caption{CUBIC vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-1}
		\end{minipage}\hfill
		\begin{minipage}{0.5\paperwidth}
			\centering
			\includegraphics[width=0.9\textwidth]{figs/dctcp-bic-8-0.png}
			\caption{DCTCP vs BIC with 8ms latency and 0\% loss}
			\label{fig:fair-2}
		\end{minipage}
	}
\end{figure}

CTCP is the only algorithm that responds excessively to others. Alone, it
reaches the full gigabit speed; but if any other algorithm runs at the same
time (i.e. a normal test is performed), CTCP's speed consistently drops. It
then reaches around 55 megabits ($\pm$5), while the other algorithms reach at %TODO: check 55+-5
least 800 with good network conditions (8 milliseconds latency, 0\% loss).
See figure \ref{fig:ctcp-1} for an example of BIC vs. CTCP.

\begin{figure}[H]
	\centering
		\includegraphics[scale=0.4]{figs/ctcp-bic-8-0.png}
		\caption{CTCP vs. BIC with 8ms latency and 0\% loss}
	\label{fig:ctcp-1}
\end{figure}

Additionally, CTCP also lowers its speed in response to the simulated network
circumstances: with 1.2\% packet loss and 8 milliseconds latency, it will reach
around 25 megabits, and with 290 milliseconds latency and no loss, it reaches
around 1.5 megabits. In neither case does it reach the original 55 megabits.

BBR is unique in that it barely responds to packet loss but only to delay: even
with 1.2\% loss, as long as the delay is low (8 milliseconds in our case), it
reaches 850$\pm$20 megabits per second. Other algorithms will reach much lower
speeds (24 megabits is not uncommon), but we see no signs of BBR being the
cause of that. In tests with BBR nor CTCP, speeds are not higher than when BBR
is one of the algorithms being tested.

The full set of results is available online\cite{git-lsproj}.

\subsection{Traffic Management}
We propose two possible solutions to manage traffic from a multitenant datacenter to customers when unfair TCP congestion control algorithms are used: Traffic shaping based Monitoring and detection of unaccepted TCP congestion algorithms in the network.

\subsubsection{Traffic shaping based Monitoring}
With the use of a hypothetical TCP congestion control algorithm that is unfair to other TCP congestion control algorithms, the bandwidth allocation to one customer can be higher to other customers using the default congestion control algorithms. For example: when a TCP congestion control algorithm shows an aggressive increase and keeps pushing data over the network, other fair TCP congestion algorithms will reduce the sending rate of the data transfer as packet loss occurs. One important condition for this unfair congestion control algorithm is that it should not be agressive such that it behaves in a Denial of Service manner.

To solve this issue, we propose a monitoring solution which monitors the bandwidth usage per server and controls the priority for the traffic sent by a server.The solution uses traffic shaping to control the priority of the traffic per server:
\begin{enumerate}
	\item{The monitoring solution detects unfair bandwidth allocation on a congested link. It detects a congested link by checking for delay and packet loss in network segments.}
	\item{The monitoring solution calculates per server whether its bandwidth usage deviates more than $\alpha$ from the average bandwidth usage. $\alpha$ is deviated bandwidth and is set according to an acceptable deviation value. The following formula is used to calculate the standard deviation: $s_{i} = \sqrt{\frac{\sum_{i=1}^N (x_i - \overline{x})^2}{N-1} }$. $N$ is the amount of active servers on the link. $\overline{x}$ is the average amount of bandwidth for all active servers on the link. $x_i$ is the bandwidth of the current server and $s_{i}$ shows the standard deviation in bandwidth for server $i$. If $s_{i}$ is a negative value, no priority change will be applied as the bandwidth usage is lower than average for the server. If $s_{i} <= \alpha$, then no priority change will be applied. If $s_{i} > \alpha$, the priority will be changed to a lower value.}
	\item{The monitoring solution provides information of the servers where the priority will be set to a lower value to the traffic shaping controller.}
	\item{The traffic shaping controller sets the priority for the servers to a lower value, so that the traffic coming from the servers that use $\alpha$ more than average bandwidth has less preference than traffic closer to the average bandwidth and lower.}
	\item{The monitoring solution periodically checks if congestion and delay are not detected. If both are not detected, the traffic shaping policy is reversed again as more bandwidth is available to allocate to customers.}
\end{enumerate}

Todo: Not sure about last enumerate item.

\subsubsection{Detection of unaccepted TCP congestion control algorithms}

\section{Discussion}\label{sec:discussion}

\section{Conclusion}\label{sec:conclusion}
To do: Monitoring $\alpha$ value research. future work.

% For reference, from the proposal:

Our main research question is:
{\it How can a datacenter provide an reliable and performant environment for
customers when multiple TCP congestion control algorithms are used in the
network?}

\vspace{0.5cm}

This main research question divides in five sub-questions:

\begin{enumerate}
	\item What TCP congestion algorithms are most commonly implemented?
	\item How do these TCP congestion algorithms work?
	\item How do these TCP congestion algorithms differ in behavior?
	\item What is the impact on the availability and service for customers when using different, competing congestion algorithms?
	\item What measures could be taken to split bandwidth equally among tenants?
\end{enumerate}

\printbibliography

\appendix
\section{Python}
\label{appendix:python}
SCRIPT

\end{document}